### 2025-06-01 paper/draft.tex
Full pass over the paper

\documentclass[10pt,leqno]{amsart}
\usepackage{graphicx}
\baselineskip=16pt
\usepackage{indentfirst,csquotes}

\topmargin= .5cm
\textheight= 20cm
\textwidth= 32cc
\baselineskip=16pt

\evensidemargin= .9cm
\oddsidemargin= .9cm

\usepackage{datetime}
\usepackage{wrapfig}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{siunitx}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{array,tabularx}
\usepackage{chngcntr}
\usepackage{afterpage}
\usepackage{ulem}
\usepackage{hyperref}
\usepackage{dirtytalk}
\usepackage{algorithm2e}
\usepackage{enumitem,amssymb}
\usepackage{pifont}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{formal-grammar}
\usepackage{varwidth}
\usepackage{fmtcount}
\usepackage{tikz}
\usepackage{array}
\usepackage[framemethod=tikz]{mdframed} % Allows defining custom boxed/framed environments
\usepackage{cleveref}
\usepackage{microtype}
\usepackage{dirtytalk}

\usetikzlibrary{trees}
\usetikzlibrary{fit}
\usetikzlibrary{shapes}
\usetikzlibrary{arrows.meta, positioning, shadows}
\usetikzlibrary{chains,shapes.multipart}
\usetikzlibrary{external}
\usetikzlibrary{matrix}

\tikzstyle{rect} = [rectangle,fill=white, text centered]
\tikzstyle{database} = [draw, cylinder, shape border rotate = 90, aspect = 0.2]
\tikzstyle{thick-arrow} = [->, thick]
\tikzstyle{arrow} = [thick,->,>=stealth]
\tikzstyle{arrow-small} = [->,>=stealth]
\tikzstyle{simple-rect} = [rectangle, text centered, draw = black, inner sep=4mm]

\tikzset{
    doc/.style={draw, minimum height=4em, minimum width=3em, 
                fill=white, 
                double copy shadow={shadow xshift=4pt, 
                             shadow yshift=4pt, fill=white, draw}}
}

\tikzset{
    dcs/.style = {double copy shadow},
}

\newtheorem{theorem}{Theorem}[]
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}

\hypersetup{
    colorlinks=true,
    filecolor=magenta,      
    urlcolor=blue,
}

\hypersetup{linkcolor=black}

\newcommand{\src}[1]{\texttt{#1}}

\lstset{basicstyle=\footnotesize\ttfamily,breaklines=true, captionpos=b}

\mdfdefinestyle{commandline}{
	leftmargin=10pt,
	rightmargin=10pt,
	innerleftmargin=15pt,
	middlelinecolor=black!50!white,
	middlelinewidth=2pt,
	frametitlerule=false,
	backgroundcolor=black!5!white,
	frametitle={LLM Output},
	frametitlefont={\normalfont\sffamily\color{white}\hspace{-1em}},
	frametitlebackgroundcolor=black!50!white,
	nobreak,
	singleextra={%
    }
}

% Define a custom environment for command-line snapshots
\newenvironment{commandline}{
	\medskip
	\begin{mdframed}[style=commandline]
}{
	\end{mdframed}
	\medskip
}

\mdfdefinestyle{warning}{
	topline=false, bottomline=false,
	leftline=false, rightline=false,
	nobreak,
	singleextra={%
		\draw(P-|O)++(-0.5em,0)node(tmp1){};
		\draw(P-|O)++(0.5em,0)node(tmp2){};
		\fill[black,rotate around={45:(P-|O)}](tmp1)rectangle(tmp2);
		\node at(P-|O){\color{white}\scriptsize\bf ?};
		\draw[very thick](P-|O)++(0,-1em)--(O);%--(O-|P);
	}
}

% Define a custom environment for warning text
\newenvironment{prompt}[1][Prompt:]{ % Set the default warning to "Warning:"
	\medskip
	\begin{mdframed}[style=warning]
		\noindent{\textbf{#1}}
}{
	\end{mdframed}
}

\title{Automatic Disease Diagnosis Using Large Language Models and Answer Set Programming}
\author{
	Ioanna Gemou
	\and
	Evangelos Lamprou
}

\begin{document}

\maketitle
\begin{abstract}
The development of automatic disease prediction systems is critically important for improving human health and supporting physicians in accurate diagnosis and treatment. Accurate diagnosis is essential for timely intervention and effective therapy, and it can also help prevent further complications and burdensome medical procedures.
Symbolic artificial intelligence has previously been applied in the medical domain \cite{Alviano_2020}. However, its adoption has been limited due to the complex knowledge bases required for such systems to operate.

The proposed method, which combines Large Language Models (LLMs) with Answer Set Programming (ASP), offers a powerful approach by providing a reliable and efficient means of generating a knowledge base that can then be used for disease prediction.
The novelty of our method lies in the combination of these two paradigms. LLMs are capable of encoding medical literature into ASP code with remarkable accuracy and deep comprehension, while ASP contributes structured reasoning and explainability. This combination enables the development of a robust disease prediction system that integrates the \textbf{accuracy and intelligence of language models} with the \textbf{explainability and efficiency of Answer Set Programming}.
Our method demonstrates promising results on the patient dataset to which it was applied.
\end{abstract}

\section{Related work}

\subsection{Large Language Models}
Large language models (LLMs) \cite{zhao2023survey} are advanced computational systems 
designed to understand and generate human language. 
Developed using cutting-edge machine learning techniques, 
these models can produce text that closely mirrors human linguistic patterns.
Modern architectures such as \textit{GPT-4} \cite{openai2023gpt4}, 
are trained on massive corpora spanning diverse domains, 
enabling them to acquire broad, cross-disciplinary knowledge. 
LLMs use deep learning algorithms \cite{Sarker2021} 
and specifically transformer networks \cite{Dosovitskiy2020} 
to analyze and understand the relationships between words, phrases, and sentences. 
By capturing patterns and structures in the training data, 
they can generate coherent and relevant responses based on given prompts.

The term “large” refers to the billions of parameters 
that allow these models to generate context-aware responses. 
The basic premise of a language model is its ability to predict 
the next word (referred to as a token) based on the text it has observed so far, 
according to the context.

% The following examples illustrate how LLMs have supported various applications \cite{Paranjape2023}.

LLMs have demonstrated strong performance across a variety of natural language tasks, 
including question answering \cite{Brown2020}, summarization \cite{Zhang2020}, 
dialogue generation \cite{Thoppilan2022lamda}, and code synthesis \cite{Chen2021}. 
In specialized domains such as biomedicine, domain-adapted LLMs have been employed 
for clinical decision support \cite{Singhal2023} and 
biomedical knowledge retrieval \cite{Luo2022}.
Among their capabilities, LLMs excel at transforming text from one representation to another, 
making them powerful tools for a wide range of natural language processing tasks. 
This characteristic is particurarly relevant to the present work, 
which focuses on the effective encoding of medical literature.

\subsection{Challenges in applying LLMs}

LLMs are not without errors \cite{Raj2023, Ruis2023}. 
They can occasionally produce incorrect or biased responses 
that reflect the biases inherent in the training data. 
There have been efforts to mitigate these issues 
through data preprocessing techniques and fine-tuning \cite{Dodge2021}.

Ongoing research seeks to improve the factual accuracy, safety, and interpretability 
of these models, enabling their deployment in high-stakes environments \cite{Ganguli2022}.

\subsection{Answer Set Programming}

Answer Set Programming (ASP) \cite{Eiter2009} is a declarative problem-solving paradigm 
rooted in logic programming and non-monotonic reasoning \cite{Brewka2011}. 
The formal semantics of stable models and the foundational ASP language 
were first introduced by Gelfond and Lifschitz \cite{Gelfond2000, gel88}.

Programming with this approach is done in a family of languages 
sometimes referred to as \textit{AnsProlog} \cite{Gelfond2002}.

The central idea of Answer Set Programming is to represent a problem 
declaratively using logical rules, facts, and constraints. 
These elements collectively form a program that describes the conditions any solution must satisfy. 
Instead of writing an algorithm to solve the problem, 
the programmer encodes the specification of valid solutions. 
An ASP solver then processes this program to compute one or more stable models (also known as answer sets),
each of which corresponds to a solution that satisfies all the given rules and constraints.

In traditional programming, transitioning from a problem to a solution 
requires the programmer's thorough understanding of the given problem, 
followed by the creation of a program that will produce the correct output 
when provided with an instance of the problem, which will then be interpreted as the solution. 
In Answer Set Programming, the process of deriving solutions from a problem specification 
typically follows a sequence of well-defined steps \cite{Gebser2013}, 
as illustrated in \cref{fig:asp-solving}.

\begin{figure}[htb]
    \begin{center}
\begin{tikzpicture}[node distance=4cm]
        \node [simple-rect] (problem)  at (0,0) {Problem};
        \node [simple-rect, below of=problem] (program) {Program};

        \node [simple-rect, right of=program] (grounder) {Grounder};
        \node [simple-rect, right of=grounder] (solver) {Solver};

        \node [simple-rect, right of=solver] (stable-models) {Stable Models};

        \node [simple-rect, above of=stable-models] (solution) {Solution};

        \draw [arrow] (problem) -- node[right] {\textbf{Modelling}} (program);
        \draw [arrow] (stable-models) -- node[left] {Interpreting} (solution);

        \draw [arrow-small] (program) -- node[above] {} (grounder);
        \draw [arrow-small] (grounder) -- node[above] {\footnotesize Grounding} (solver);
        \draw [arrow-small] (solver) -- node[above] {} (stable-models);

        \node[draw, dotted, fit=(grounder) (solver), inner sep=4mm, label=above:{Solving}] {};
\end{tikzpicture}
    \end{center}
		\caption{Solving a problem using ASP.}
    \label{fig:asp-solving}
\end{figure}

\begin{itemize}
    \item \textbf{Modeling}: The problem is modeled using ASP (Answer Set Programming) language.
    \item \textbf{Grounding}: A grounder (e.g., gringo) converts the program into a set of basic rules and facts.
    \item \textbf{Solving}: A solver (e.g., clasp) finds a solution to the problem by computing the set of stable models (answer sets).
\end{itemize}

Several solvers for ASP have been developed, such as DLV \cite{Xia2020}. 
This work uses the ASP system clingo \cite{Gebser2014}, 
which combines the grounder gringo with the clasp solver \cite{Holldobler2014} 
into a single application, also providing a powerful Python API for integrating the solver into other applications.

\subsection{ASP for disease diagnosis}
ASP offers a powerful method for representing medical knowledge in a declarative and logical manner. 
Some common ASP methods used in the representation of medical knowledge \cite{Vinarti2019}:

\begin{itemize}
    \item \textbf{Rule-Based Representation}: ASP allows for the representation of medical knowledge as rules in the form of logical statements. These rules define relationships between medical entities such as symptoms, diseases, treatments, and patient data. For example, a rule could state that if a patient exhibits specific symptoms, it indicates the presence or absence of certain diseases.
    \item \textbf{Knowledge Base Construction}: ASP provides a framework for constructing a knowledge base that captures medical knowledge. The knowledge base consists of facts, rules, and constraints and may capture a wide range of medical information, including symptom-disease associations, diagnostic criteria, treatment protocols, and patient-specific data.
\end{itemize}

Earlier methods address complex queries over biomedical ontologies 
and databases by extracting relevant information from structured 
knowledge sources \cite{Erdem2011}. Applied in drug discovery, 
these techniques integrate biological databases with concise, 
interpretable explanations for complex queries related to drug development.

\section{Methodology}

\subsection{Architecture}

\begin{figure}[!h]
\begin{center}
\begin{tikzpicture}[node distance = 5cm, minimum width=2cm, minimum height=1cm]
    \node [doc, label=above:{}] (med-bib) {Medical Bibliography};
    \node [rect, draw, right=3cm of med-bib, fill=blue!50] (llm) {Large Language Model};
    \node [rect, draw, above= 1cm of med-bib, fill=orange!50] (prompt) {Prompt};
    \node [database, below left= 3cm of med-bib, fill=green!30] (knowledge-base) {Knowledge Base};
    \node [rect, draw, right of=knowledge-base] (solver) {ASP Solver};
    \node [rect, draw, below=1cm of solver] (patient) {\textit{Patient Data}};
    \node [rect, right of=solver] (diagnosis) {\textbf{Diagnosis}};

    \draw[thick-arrow] (med-bib) edge [] node [above] {} (llm);
    \draw[thick-arrow] (prompt) edge [in=90, out=0, looseness=1] node [above, sloped] {\textit{\say{please convert the following into ASP code...}}} (llm);
    \draw[thick-arrow] (llm) edge [in=90, out=270, looseness=1] node [below] {ASP Program} (knowledge-base);
    \draw[thick-arrow] (knowledge-base) edge [] node [above] {} (solver);
    \draw[thick-arrow] (solver) edge [] node [above] {} (diagnosis);
    \draw[thick-arrow] (patient) edge [in=270, out=180, looseness=1] node [above] {} (knowledge-base);

\end{tikzpicture}
\end{center}
\caption{The diagnostic process begins by transforming medical literature 
    into an ASP program using a LLM. 
    An ASP solver then executes the program and produces a final diagnosis 
    based on each patient’s data.}
    \label{fig:architecture}
\end{figure}

The goal is to extract a structured knowledge base from unstructured medical text \cite{ré2014feature}.
\Cref{fig:architecture} illustrates the proposed architecture, which follows the key steps below:

\begin{itemize}
    \item Collect medical literature relevant to the diseases or conditions of interest.
    \item Design a suitable prompt to convert the medical literature into ASP code.
    \item Incorporate the ASP code generated by the LLM into the knowledge base.
    \item Add relevant patient data—such as symptoms and examination results—to the knowledge base.
    \item Run an ASP solver to reason over the final knowledge base and produce the diagnosis or set of possible diagnoses as answer sets.
\end{itemize}

\subsection{Prompt Construction}

While large language models demonstrate strong instruction-following capabilities, 
their reliability remains limited in practice \cite{bowman2023things}. 
To address this, a relatively recent methodology known as prompt engineering 
has emerged \cite{liu2021pretrain, promptingguidePromptEngineering}. 
In this work, prompts are carefully crafted to guide the model in 
converting medical literature into ASP code suitable for execution 
by a solver (\cref{fig:bad-prompt}, \cref{fig:good-prompt}).
\footnote{The example of a poorly constructed prompt is included for educational purposes, to highlight common pitfalls in prompt design.}

% Ioanna: I commented out the bad prompt
\begin{figure}[!h]
\centering
\begin{prompt}
\textit{Medical Text...}

The paragraph above lists common symptoms of pneumonia.
Write a Clingo script that diagnoses pneumonia based on these symptoms.
\end{prompt}

\begin{commandline}
\begin{verbatim}

symptom(cough).
symptom(cough_with_mucus).
symptom(chest_pain).
symptom(headache).
...

diagnosis(pneumonia) :- symptom(cough), 
                        symptom(chest_pain), 
                        symptom(headache), 
                        symptom(arthralgia),
                        ...
\end{verbatim}
\end{commandline}

\caption{A simple prompt that does not yield satisfactory results.}
\label{fig:bad-prompt}
\end{figure}

\begin{figure}[!h]
\begin{prompt}
\textit{Medical Text...}

The paragraph above lists common symptoms of pneumonia.
Write a Clingo script that diagnoses pneumonia based on these symptoms.

In the diagnosis rule, use a structure like:

\src{diagnosis(pneumonia) :- has(symptom(x)), has(symptom(y)) ...}

Include alternative diagnoses that share overlapping symptoms with pneumonia.
Add rules that link one symptom to another (e.g., symptom propagation or dependency).

\end{prompt}

\begin{commandline}
\scriptsize
\begin{verbatim}
symptom(cough).
symptom(cough_with_mucus).
...

linked_symptom(cough_with_mucus, wheezing).
linked_symptom(cough_with_mucus, shortness_of_breath).

...

has(symptom(Symptom2)) :- has(symptom(Symptom1)), 
              linked_symptom(Symptom1, Symptom2).

diagnosis(pneumonia) :- has(symptom(cough)),
                        has(symptom(cough_with_mucus)),
...

diagnosis(pneumonia) :- has(symptom(cough_with_mucus)),
                        has(symptom(chest_pain)),
...
\end{verbatim}
\end{commandline}
\caption{An effective prompt that produces accurate and well-structured output.}
\label{fig:good-prompt}
\end{figure}

\subsection{Construction of a Logical Program}

The knowledge base is initialized with a set of symptoms and medical indicators represented as:
\begin{align}
    symptom(s). & \; s \in \{ \text{cough, chest pain, rash}, \dots \} \\
    indicator(i). & \; i \in \{ \text{low MCV, high TIBC}, \dots \}
\end{align}

A patient exhibiting a specific symptom or testing positive for a particular indicator is represented as:
\begin{equation}
    has(x). \; x \in S \cup I
\end{equation}

Logical rules used for inferring diagnoses follow this general form:
\begin{align}
    diagnosis(d_1) & \longleftarrow has(x_1) \land has(x_2) \land \dots \\
    diagnosis(d_1) & \longleftarrow has(x_1) \land has(x_3) \land \dots \\
    diagnosis(d_2) & \longleftarrow has(x_3) \land has(x_4) \land \dots
\end{align}

Correlations between certain symptoms are also encoded in the knowledge base. 
For example, clinical observations have reported links between skin conditions and headaches \cite{migraine-hives}. 
In some cases, the presence of one symptom may lead to the appearance of another. 
Such relationships are captured using rules of the form:
\begin{align}
    has(symptom(S2)) \longleftarrow has(symptom(S1)) \land linked\_symptom(S1, S2).
\end{align}

Example entries for such relationships include:
\begin{align}
    linked\_symptom(grunting, chest\_retractions).\\
    linked\_symptom(nasal\_flaring, chest\_retractions).
\end{align}

Since patients may not exhibit \textit{all} known symptoms, 
the system must support reasoning under partial information. 
To allow flexible reasoning, a choice rule introduces possible symptoms 
into the knowledge base when patient data is incomplete:
\begin{equation}
    \{ add(symptom(S)) : symptom(S) \}.
\end{equation}

To ensure that the solver identifies at least one diagnosis, 
the following integrity constraint is imposed:
\begin{equation}
    \bot \longleftarrow \sim diagnosis(\_)
\end{equation}

Finally, the solver minimizes assumptions by penalizing added symptoms, 
favoring diagnoses based on known patient information:
\begin{equation}
    \#minimize \{ 1, S : add(S) \}
\end{equation}

This formulation guarantees that at least one diagnosis is produced, 
relying as much as possible on available clinical data rather than hypothetical assumptions.

\subsection{Explainability of Diagnoses}

Symbolic artificial intelligence offers a significant advantage 
over subsymbolic methods in terms of explainability. 
In symbolic systems, the relationships between facts and conclusions are explicitly defined, 
enabling transparent reasoning. However, ASP solvers such as \textit{Clingo} do not, 
at the time of writing,\footnote{At the time of writing this paper.} 
provide intuitive traces of their reasoning processes. 
Prior work has explored methods for associating logical propositions 
with their justifications \cite{cabalar2014causal}. For example, 
the logic program shown below is accompanied by a causal graph 
in \cref{fig:causal-g}, which illustrates the connections 
between rules and their inferred outcomes. 
This structure supports the interpretability of conclusions derived by a symbolic logic solver.
\begin{equation}
\begin{aligned}[b]
    l :& punish \longleftarrow drive, drunk \\
    m :& punish \longleftarrow resist \\
    e :& prison \longleftarrow punish \\
    d :& drive \\
    k :& drunk \\
    r :& resist \\
\end{aligned}
\label{eq:example-cg}
\end{equation}

\begin{figure}
    \centering
    \includegraphics{assets/causal_g.png}
    \caption{Causal graph of logic program (\cref{eq:example-cg}).}
    \label{fig:causal-g}
\end{figure}

The tool \textit{xclingo}\footnote{\url{https://github.com/bramucas/xclingo2}} \cite{Cabalar_2020} 
allows the presentation of justifications regarding the results given by the solver \textit{Clingo}. 
This is achieved by adding a \src{trace} corresponding to the rules of our logic program. 
roviding a clear explanation of how the system reached specific conclusions can help identify potential 
errors in the creation of the knowledge base and provide the system user with a rationale for the final outcome. 
This is particularly important when the application is used in medical contexts.

\section{Results}

Evaluation was conducted using publicly available patient data.\footnote{\url{https://www.kaggle.com/datasets/itachi9604/disease-symptom-description-dataset?select=dataset.csv}} 
Each patient's symptoms were provided as input to the system. 
The knowledge base combines all rules derived from the encoded medical literature. 
A prediction is considered successful when the disease identified by the ASP solver matches the patient’s actual diagnosis.
\begin{table}[h]
    \centering
    \begin{tabular}{lcc}
    \hline
    Diseases & Size of Knowledge Base & Accuracy \\
    \hline
    \hline
    Chicken Pox      &  66  & 95\% \\
    Pneumonia        &  75  & 100\% \\
    Common Cold    &   44 & 100\% \\
    \hline
    \end{tabular}
    \caption{Accuracy of the method on selected diseases.}
\end{table}

The model achieves high accuracy when a sufficiently 
large set of rulesis included in the knowledge base. 
The size of the knowledge base is defined as the number 
of logical terms in the ASP program associated with each disease.

The system also incorporates an explanation mechanism using the \src{xclingo} tool,
which generates a graph visualizing the logical associations used to derive the diagnosis.

\begin{lstlisting}[caption={A part of the explanation for the chickenpox diagnosis. 
    The decision tree illustrates the symptom associations and outlines 
    the reasoning process used by the system to reach its conclusion.},
label=lst:terms]{Name}
  *
  |__diagnosis(chickenpox)
  |  |__has(symptom(itching));has(symptom(itching))
  |  |__has(symptom(fatigue));has(symptom(fatigue));has(symptom(fatigue));has(symptom(fatigue));has(symptom(fatigue))
  |  |__has(symptom(lethargy))
  |  |  |__has(symptom(fatigue));has(symptom(fatigue));has(symptom(fatigue));has(symptom(fatigue));has(symptom(fatigue))
  |  |  |__linked_symptom(fatigue,lethargy)
  |  |__has(symptom(high_fever))
  |  |  |__has(symptom(mild_fever));has(symptom(mild_fever))
  |  |  |  |__has(symptom(loss_of_appetite));has(symptom(loss_of_appetite))
  |  |  |  |__linked_symptom(loss_of_appetite,mild_fever)
  |  |  |__linked_symptom(mild_fever,high_fever)
  |  |__has(symptom(loss_of_appetite));has(symptom(loss_of_appetite))
  |  |__has(symptom(mild_fever));has(symptom(mild_fever))
  |  |  |__has(symptom(loss_of_appetite));has(symptom(loss_of_appetite))
  |  |  |__linked_symptom(loss_of_appetite,mild_fever)
  |  |__has(symptom(swelled_lymph_nodes))
  |  |  
  ...
\end{lstlisting}

\section{Conclusions}

This work presents a method for automatic disease prediction 
that encodes medical literature and transforms it into ASP code using an LLM. 
It describes the roles of both LLMs and ASP, emphasizing how each contributes to the overall system. 
The system architecture is outlined, along with implementation details for constructing effective prompts 
and structuring rules within the ASP knowledge base.
The method includes predictions for a set of diseases based on the constructed knowledge bases. 
Results show that this approach achieves high-accuracy predictions and provides interpretable explanations.

Future work includes integrating retrieval-augmented generation (RAG) to improve the factual accuracy 
and reliability of the ASP code, and fine-tuning the LLM on clinical reasoning tasks to enhance rule quality. 
Expanding the approach to cover additional diseases is also a key direction.
\newpage
\bibliographystyle{unsrt}
\bibliography{bib}

\end{document}