### 2025-12-30 paper/draft.tex
Prepare for arxiv
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

\documentclass[11pt,leqno]{amsart}
\usepackage{graphicx}
\baselineskip=16pt
\usepackage{indentfirst,csquotes}
\usepackage[margin=1in]{geometry}

% \topmargin= .5cm
% \textheight= 20cm
% \textwidth= 32cc
% \baselineskip=16pt

% \evensidemargin= .9cm
% \oddsidemargin= .9cm

\usepackage{datetime}
\usepackage{wrapfig}
\usepackage{enumerate}
\usepackage{caption}
\usepackage{siunitx}
\usepackage{subcaption}
\usepackage{array,tabularx}
\usepackage{chngcntr}
\usepackage{afterpage}
\usepackage{ulem}
\usepackage{hyperref}
\usepackage{dirtytalk}
\usepackage{algorithm2e}
\usepackage{enumitem,amssymb}
\usepackage{pifont}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{formal-grammar}
\usepackage{varwidth}
\usepackage{fmtcount}
\usepackage{tikz}
\usepackage[framemethod=tikz]{mdframed} % Allows defining custom boxed/framed environments
\usepackage{cleveref}
\usepackage{microtype}
\usepackage{booktabs}

\usetikzlibrary{trees}
\usetikzlibrary{fit}
\usetikzlibrary{shapes}
\usetikzlibrary{arrows.meta, positioning, shadows}
\usetikzlibrary{chains,shapes.multipart}
\usetikzlibrary{external}
\usetikzlibrary{matrix}

\newcommand{\eg}{{e.g.}, }
\newcommand{\ie}{{i.e.}, }
\newcommand{\etc}{{etc.}\xspace}
\newcommand{\vs}{{vs.} }
\newcommand{\cf}[1]{(\emph{Cf}.\S\ref{#1})}
\newcommand{\sx}[1]{(\cref{#1})}

\tikzstyle{rect} = [rectangle,fill=white, text centered]
\tikzstyle{database} = [draw, cylinder, shape border rotate = 90, aspect = 0.2]
\tikzstyle{thick-arrow} = [->, thick]
\tikzstyle{arrow} = [thick,->,>=stealth]
\tikzstyle{arrow-small} = [->,>=stealth]
\tikzstyle{simple-rect} = [rectangle, text centered, draw = black, inner sep=4mm]

\tikzset{
    doc/.style={draw, minimum height=4em, minimum width=3em, 
                fill=white, 
                double copy shadow={shadow xshift=4pt, 
                             shadow yshift=4pt, fill=white, draw}}
}

\tikzset{
    dcs/.style = {double copy shadow},
}

\newtheorem{theorem}{Theorem}[]
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}

\hypersetup{
    colorlinks=true,
    filecolor=magenta,      
		citecolor=black,
    urlcolor=blue,
}

\hypersetup{linkcolor=black}

\newcommand{\src}[1]{\texttt{#1}}

\lstset{basicstyle=\footnotesize\ttfamily,breaklines=true, captionpos=b}

\mdfdefinestyle{commandline}{
	leftmargin=10pt,
	rightmargin=10pt,
	innerleftmargin=15pt,
	middlelinecolor=black!50!white,
	middlelinewidth=2pt,
	frametitlerule=false,
	backgroundcolor=black!5!white,
	frametitle={LLM Output},
	frametitlefont={\normalfont\sffamily\color{white}\hspace{-1em}},
	frametitlebackgroundcolor=black!50!white,
	nobreak,
	singleextra={%
    }
}

% Define a custom environment for command-line snapshots
\newenvironment{commandline}{
	\medskip
	\begin{mdframed}[style=commandline]
}{
	\end{mdframed}
	\medskip
}

\mdfdefinestyle{warning}{
	topline=false, bottomline=false,
	leftline=false, rightline=false,
	nobreak,
	singleextra={%
		\draw(P-|O)++(-0.5em,0)node(tmp1){};
		\draw(P-|O)++(0.5em,0)node(tmp2){};
		\fill[black,rotate around={45:(P-|O)}](tmp1)rectangle(tmp2);
		\node at(P-|O){\color{white}\scriptsize\bf !};
		\draw[very thick](P-|O)++(0,-1em)--(O);%--(O-|P);
	}
}

% Define a custom environment for warning text
\newenvironment{prompt}[1][Prompt:]{ % Set the default warning to "Warning:"
	\medskip
	\begin{mdframed}[style=warning]
		\noindent{\textbf{#1}}
}{
	\end{mdframed}
}

\newcommand{\sys}{\textsc{McCoy}\xspace}

\makeatletter
\def\@settitle{%
  \begin{center}%
    \normalfont\LARGE\bfseries
    \@title\par
  \end{center}%
}
\makeatother

\title[A PoC for Explainable Disease Diagnosis Using LLMs \& ASP]{A Proof-of-Concept for Explainable Disease Diagnosis\\ Using Large Language Models and\\ Answer Set Programming}

\author[Ioanna Gemou]{Ioanna Gemou}
\author[Evangelos Lamprou]{Evangelos Lamprou}
\thanks{Originally submitted as part of a university research competition on 1st June 2023. Both authors were students at the University of Patras, Greece, at the time.}

\begin{document}

\maketitle

\begin{abstract}
Accurate disease prediction is vital for timely intervention, 
effective treatment, and reducing medical complications. 
While symbolic AI has been applied in healthcare, 
its adoption remains limited due to the effort required for constructing high-quality knowledge bases.
This work introduces \sys, a framework that combines Large Language Models (LLMs) with Answer Set Programming (ASP) to overcome this barrier. 
\sys orchestrates an LLM to translate medical literature into ASP code, combines it with patient data, and processes it using an ASP solver to arrive at the final diagnosis.
This integration yields a robust, interpretable prediction framework that leverages the strengths of both paradigms. 
Preliminary results show \sys has strong performance on small-scale disease diagnosis tasks.
% \sys accurately predicts diseases based on symptoms derived from medical literature.
\end{abstract}

\section{Introduction}
Getting an accurate diagnosis quickly can make all the difference in patient outcomes, 
but building automated systems that practitioners can trust is still a major challenge~\cite{Vinarti2019}.
Symbolic AI, and in particular Answer Set Programming (ASP), provides clear logical reasoning~\cite{Brewka2011,Gelfond_Kahl_2014}
and has seen success in the medical domain~\cite{Alviano_2020}.
However, its adoption has been limited because constructing and maintaining large, 
domain-specific knowledge bases is very labor-intensive.
On the other hand, Large Language Models (LLMs) have shown strong results on tasks
such as medical question answering and clinical note summarization \cite{Singhal2022LargeLM,Singhal2023,Zheng2024LLMmed}, 
but their tendency to produce inconsistent or opaque outputs makes them risky for clinical use~\cite{aljohani2025comprehensivesurveytrustworthinesslarge,rudin2019stopexplainingblackbox,Bommasani2021}.

There is an acute need for clinical support systems that are both interpretable and widely applicable.
\sys takes a step in that direction by combining the strengths of LLMs and ASP. 
Instead of requiring experts to hand-craft large knowledge bases, 
\sys automatically converts medical literature into ASP rules. 
These rules are then enriched with patient symptoms and clinical indicators, 
and passed through an ASP solver to produce diagnostic suggestions together with logical justifications.
On public datasets, \sys reaches 95–100\% predictive accuracy on selected diseases, 
while giving transparent explanations for each diagnosis.

% Concretely, this work:
% \begin{itemize}
%     \item Introduces a framework that combines the flexibility of LLMs with the explanability of ASP for disease diagnosis.
%     \item Provides solver-based explanations showing how symptoms lead to rules and then to diagnoses.
%     \item Demonstrates accurate results on real patient data while keeping the reasoning fully transparent.
% \end{itemize}

The paper begins with background on LLMs and ASP (\cref{sec:background}), 
then details the \sys framework (\cref{sec:methodology}), 
evaluates the approach on patient data (\cref{sec:results}), 
and concludes with a discussion of limitations and future directions (\cref{sec:conclusion}).

\section{Background}\label{sec:background}

\paragraph{\textbf{Large Language Models}}
LLMs are advanced computational systems 
designed to understand and generate human language~\cite{zhao2023survey}. 
These models can produce text that closely mirrors human linguistic patterns and have emergent reasoning capabilities~\cite{Paranjape2023}.
Recent models such as \textit{GPT-5}~\cite{gpt5}
have been trained on massive corpora spanning diverse domains, 
enabling them to acquire broad, cross-disciplinary knowledge. 
To give an intuition of the way they operate, LLMs are deep neural networks that have converged to transformer networks~\cite{Vaswani2017},
which are trained using self-supervised learning~\cite{Gui2024surveySSL}.
During training, these models refine an internal representation that enables them to analyze and understand the relationships between words, phrases, and sentences. 
By capturing patterns and structures in the training data, 
they can then be probed to generate coherent and relevant responses based on given prompts.
The term ``large" often refers to models with a large number of parameters (often in the billions)
that allow these models to build high-quality representations of language.
At a very basic level, an LLM is able to predict the next word (referred to as a token) based on its preceding context.

% The following examples illustrate how LLMs have supported various applications \cite{Paranjape2023}.

LLMs have demonstrated strong performance across a variety of natural language tasks, 
including question answering \cite{Brown2020}, summarization \cite{Zhang2020}, 
dialogue generation \cite{Thoppilan2022lamda}, and program synthesis \cite{Chen2021}. 
In specialized domains such as biomedicine, domain-adapted LLMs have been employed 
for clinical decision support \cite{Singhal2023} and 
biomedical knowledge retrieval \cite{Luo2022}.
Among their capabilities, LLMs excel at transforming text from one representation to another~\cite{Vaswani2017,Li2024LLMDataProcessing}.
This characteristic is particularly relevant to the present work, 
which focuses on the effective encoding of medical literature. \\

\paragraph{\textbf{Challenges in applying LLMs}}

However, LLMs are not without errors~\cite{Raj2023, Ruis2023}. 
They can occasionally produce incorrect or nonsensical responses 
\cite{Huang2025SurveyHallucinationLLM}.
There have been efforts to mitigate these issues 
through data preprocessing techniques and fine-tuning \cite{Dodge2021}.
Ongoing research seeks to improve the factual accuracy, safety, and interpretability 
of these models, enabling their deployment in high-stakes environments \cite{Ganguli2022}.
This work specifically tackles this issue by grounding LLM outputs on real medical knowledge, and having an ASP solver provide further validation and explainability.
\\

\paragraph{\textbf{Answer Set Programming}}

Answer Set Programming (ASP) \cite{Eiter2009} is a declarative problem-solving paradigm 
rooted in logic programming and non-monotonic reasoning \cite{Brewka2011}. 
The formal semantics of stable models and the foundational ASP language 
were first introduced by Gelfond and Lifschitz \cite{gel88}.
These are implemented in a family of languages commonly known as \textit{AnsProlog} \cite{Gelfond2002}.

The central idea of ASP is to represent a problem 
declaratively using logical rules, facts, and constraints. 
These elements collectively form a program that describes the conditions any solution must satisfy. 
% Instead of writing an algorithm to solve the problem, 
% the programmer encodes the specification of valid solutions. 
Given a valid program, an ASP solver processes it to compute one or more stable models (also known as answer sets),
each of which corresponds to a solution that satisfies all the given rules and constraints.

In traditional programming, transitioning from a problem to a solution 
requires the programmer's thorough understanding of the given problem, 
followed by the creation of a program that will take as input an instance of the problem,
and produce output which will then be interpreted as the solution. 
In ASP, the process of deriving solutions from a problem specification 
typically follows a sequence of well-defined steps \cite{Gebser2013}, 
as illustrated in \cref{fig:asp-solving}. 
First, the problem is modeled using the ASP language. 
Next, a grounder (\eg gringo \cite{Gebser2014}) transforms the program into a set of basic rules and facts. 
Finally, a solver (\eg clasp \cite{Holldobler2014}) computes the stable models (answer sets), 
which correspond to the solutions of the original problem.

\begin{figure}[t]
    \begin{center}
\begin{tikzpicture}[node distance=2.9cm]
        \node [simple-rect] (problem)  at (0,0) {Problem};
        \node [simple-rect, below of=problem] (program) {Program};

        \node [simple-rect, right of=program] (grounder) {Grounder};
        \node [simple-rect, right=2cm of grounder] (solver) {Solver};

        \node [simple-rect, right of=solver] (stable-models) {Stable Models};

        \node [simple-rect, above of=stable-models] (solution) {Solution};

        \draw [arrow] (problem) -- node[right] {\textbf{Modeling}} (program);
        \draw [arrow] (stable-models) -- node[left] {Interpreting} (solution);

        \draw [arrow-small] (program) -- node[above] {} (grounder);
        \draw [arrow-small] (grounder) -- node[above] {\footnotesize Grounding} (solver);
        \draw [arrow-small] (solver) -- node[above] {} (stable-models);

        \node[draw, dotted, fit=(grounder) (solver), inner sep=2mm, label=above:{Solving}] {};
\end{tikzpicture}
    \end{center}
		\caption{\textbf{Solving a problem with ASP}. A problem is first modeled as a program, then transformed into ground rules by a grounder. A solver computes the stable models, which are interpreted as solutions.}
    \label{fig:asp-solving}
\end{figure}

Several solvers for ASP have been developed, with some seeing industrial adoption~\cite{Xia2020,Gebser2014}.
This work uses the Clingo ASP system,
which integrates the \src{gringo} grounder \cite{Gebser2014} and the \src{clasp} solver \cite{Holldobler2014} into a single pipeline.
The Clingo system also offers a powerful Python API, enabling integration into other applications.

\section{Related work}
\paragraph{\textbf{ASP for disease diagnosis}}
ASP has been used in the past to represent and reason over medical knowledge in a declarative, logic-based manner. 
Early work applied ASP to biomedical ontologies and structured databases to answer complex diagnostic 
and pharmacological queries with interpretable logical justifications \cite{Erdem2011}. 
In the clinical domain, ASP is especially effective for encoding medical rules and patient-specific data, 
enabling precise and explainable decision-making~\cite{Alviano_2020}

Two common strategies illustrate the role of ASP in medical reasoning. 
The first is \textit{rule-based representation}, where medical knowledge is expressed as logical statements 
that capture relationships among symptoms, diseases, treatments, and patient data. 
For instance, a rule might specify that if a patient exhibits certain symptoms, 
this supports or excludes the presence of particular diseases. 
The second is \textit{knowledge base construction}, where facts, rules, and constraints 
are combined into structured repositories that consistently model diagnostic criteria, 
symptom–disease associations, and therapeutic protocols in an extensible way. 
This work builds on these foundations and works towards closing the gap with fully automated knowledge-base construction.
% Together, these approaches demonstrate how ASP provides a flexible yet rigorous framework 
% for supporting medical diagnosis, offering not only accurate predictions 
% but also transparent reasoning paths that can be trusted in high-stakes settings.
\\

\paragraph{\textbf{Integrating LLMs with ASP}}

There has been work that explores the synergy between LLMs and ASP 
for enhanced diagnostic reasoning. 
These hybrid approaches aim to combine the language understanding of LLMs 
with the precision and explainability of symbolic reasoning. 
A YAML-based interface has been proposed to guide LLMs 
in extracting structured facts from natural language, 
which are then processed by ASP for logical inference \cite{alviano2024llm2asp}. 
Other work uses ASP to validate and trace misleading or inconsistent outputs 
from LLMs in medical contexts \cite{Nguyen2025}. 
Another work leverages smaller-sized LLMs and uses fine-turning to enhance their performance in generating valid ASP programs~\cite{coppolillo2024llasp}.
% Collectively, these systems assist in building reliable and clear reasoning pipelines by combining language understanding with rule-based validation.
These works share several of the same insights and motivations as \sys.

\section{Methodology} \label{sec:methodology}

\paragraph{\textbf{Architecture}}

\Cref{fig:architecture} illustrates \sys's architecture, 
which transforms unstructured medical text into a structured knowledge base. 
The process begins with the collection of medical literature relevant to the diseases or conditions of interest. 
A carefully designed prompt is then used to guide an LLM in translating this literature into ASP code, 
which is incorporated into a growing knowledge base. 
This knowledge base is further enriched with patient-specific information such as symptoms and examination results. 
An ASP solver subsequently reasons over the combined knowledge base and produces a set of plausible diagnoses.

\begin{figure}[t]
    % https://docs.google.com/drawings/d/1cr2lTIjSUxfQH54CZui8vGqLS2keOfW41i4g4uGfwhw/edit?usp=sharing
    \includegraphics[width=\textwidth]{assets/mccoy.pdf}
    \caption{\textbf{\sys overview.} The diagnostic process begins by transforming medical literature 
    into an ASP program using a LLM. 
    An ASP solver then executes the program and produces a final diagnosis 
    based on each patient’s data.}
    \label{fig:architecture}
\end{figure}

\paragraph{\textbf{Prompt Construction}}
While Large Language Models exhibit strong instruction-following capabilities, 
their reliability in high-stakes domains remains limited \cite{bowman2023things}. 
To improve robustness, a wide range of prompting strategies are employed by \sys~\cite{liu2021pretrain, promptingguidePromptEngineering}.
In \sys, carefully designed prompts are employed to guide the model in translating medical literature.
This involves a careful and accurate description of the desired output format, and specific instructions on the kind of information to encode within it.
\Cref{fig:bad-prompt} represents an example of a poorly constructed prompt that fails to yield satisfactory results,
and \cref{fig:good-prompt} shows the prompt that \sys ultimately uses to convert medical text into ASP code.

\begin{figure}[!h]
\centering
\begin{prompt}
\textit{Medical Text...}

The paragraph above lists common symptoms of pneumonia.
Write a clingo script that diagnoses pneumonia based on these symptoms.
\end{prompt}

\begin{commandline}
\footnotesize
\begin{verbatim}

symptom(cough).
symptom(cough_with_mucus).
symptom(chest_pain).
symptom(headache).
...

diagnosis(pneumonia) :- symptom(cough), 
                        symptom(chest_pain), 
                        symptom(headache), 
                        symptom(arthralgia),
                        ...
\end{verbatim}
\end{commandline}

\caption{A prompt that does not yield satisfactory results. The LLM output does not decompose the disease diagnosis into discrete rules.}
\label{fig:bad-prompt}
\end{figure}

\begin{figure}[!h]
\begin{prompt}
\textit{Medical Text...}

The paragraph above lists common symptoms of pneumonia.
Write a clingo script that diagnoses pneumonia based on these symptoms.
In the diagnosis rule, use a structure like:

\src{diagnosis(pneumonia) :- has(symptom(x)), has(symptom(y)) ...}

Include alternative diagnoses that share overlapping symptoms with pneumonia.
Add rules that link one symptom to another (\eg symptom propagation or dependency).

\end{prompt}

\begin{commandline}
\footnotesize
\begin{verbatim}
symptom(cough).
symptom(cough_with_mucus).
...

linked_symptom(cough_with_mucus, wheezing).
linked_symptom(cough_with_mucus, shortness_of_breath).

...

has(symptom(Symptom2)) :- has(symptom(Symptom1)), 
              linked_symptom(Symptom1, Symptom2).

diagnosis(pneumonia) :- has(symptom(cough)),
                        has(symptom(cough_with_mucus)),
...

diagnosis(pneumonia) :- has(symptom(cough_with_mucus)),
                        has(symptom(chest_pain)),
...
\end{verbatim}
\end{commandline}
\caption{An effective prompt that produces accurate and well-structured output.
	The LLM output is structured into multiple rules that aid explainability and can result in partial diagnoses.}
\label{fig:good-prompt}
\end{figure}

\paragraph{\textbf{Logical Program Construction}}

The knowledge base is initialized with a set of symptoms and medical indicators represented as:
\begin{align}
    symptom(s). & \; s \in \{ \text{cough, chest pain, rash}, \dots \} \\
    indicator(i). & \; i \in \{ \text{low MCV, high TIBC}, \dots \}
\end{align}
Let $S$ denote the set of symptoms and $I$ the set of indicators.
A patient exhibiting a specific symptom or testing positive for an indicator is represented as:
\begin{equation}
    has(x). \; x \in S \cup I
\end{equation}

Logical rules used for inferring diagnoses follow this general form:
\begin{align}
    diagnosis(d_1) & \longleftarrow has(x_1) \land has(x_2) \land \dots \\
    diagnosis(d_1) & \longleftarrow has(x_1) \land has(x_3) \land \dots \\
    diagnosis(d_2) & \longleftarrow has(x_3) \land has(x_4) \land \dots
\end{align}

Correlations between certain symptoms are also encoded in the knowledge base. 
For example, clinical observations have reported links between skin conditions and headaches \cite{migraine-hives}. 
In some cases, the presence of one symptom may lead to the appearance of another. 
Such relationships are captured using rules of the form:
\begin{align}
    has(symptom(S2)) \longleftarrow has(symptom(S1)) \land linked\_symptom(S1, S2).
\end{align}

Example entries for such relationships include:
\begin{align}
    linked\_symptom(grunting, chest\_retractions).\\
    linked\_symptom(nasal\_flaring, chest\_retractions).
\end{align}

Since patients may not exhibit \textit{all} known symptoms, 
\sys must support reasoning under partial information. 
To allow flexible reasoning, a choice rule introduces possible symptoms 
into the knowledge base when patient data is incomplete:
\begin{equation}
    \{ add(symptom(S)) : symptom(S) \}.
\end{equation}

The following integrity constraint imposes at least one diagnosis:
\begin{equation}
    \bot \longleftarrow \sim diagnosis(\_)
\end{equation}

Finally, the solver minimizes assumptions by penalizing added symptoms, 
favoring diagnoses based on known patient information:
\begin{equation}
    \#minimize \{ 1, S : add(symptom(S)) \}
\end{equation}

This formulation guarantees that at least one diagnosis is produced, 
relying as much as possible on available clinical data rather than hypothetical assumptions.
\\

\paragraph{\textbf{Diagnostic Explainability}}

Symbolic AI offers a significant advantage 
over sub-symbolic methods in terms of explainability. 
In symbolic systems, the relationships between facts and conclusions are explicitly defined, 
enabling transparent reasoning.
Thus, it possible to associate logical propositions with their justifications~\cite{cabalar2014causal}.
For example, 
the logic program shown in \cref{eq:example-cg} is accompanied by a causal graph 
shown in \cref{fig:causal-g}, which illustrates the connections 
between rules and their inferred outcomes. 
This structure supports the interpretability of conclusions derived by a symbolic logic solver.
The tool \src{xclingo}~\cite{Cabalar_2020} 
enables the visualization of justifications for results produced by the \src{clingo} solver. 
It does so by incorporating a trace that corresponds to the rules of the logic program. 
\sys leverages this produce such a trace, which can then help identify the reasoning path, as well as potential flaws in \sys's knowledge base and reasoning process.

% \begin{equation}
% \begin{aligned}[b]
%     l :& punish \longleftarrow drive, drunk \\
%     m :& punish \longleftarrow resist \\
%     e :& prison \longleftarrow punish \\
%     d :& drive \\
%     k :& drunk \\
%     r :& resist \\
% \end{aligned}
% \label{eq:example-cg}
% \end{equation}
% % https://docs.google.com/drawings/d/1Ing4GFH2kbniiHI5rbXzw-xcAUQAxWAJpgmg4tz1fzc/edit?usp=sharing
% \begin{figure}
%     \centering
%     \includegraphics{assets/causal_g.pdf}
%     \caption{Causal graph of logic program (\cref{eq:example-cg}).}
%     \label{fig:causal-g}
% \end{figure}
\begin{figure}[t]
    \centering
    \begin{minipage}{0.48\textwidth}
    \begin{equation}
    \begin{aligned}[b]
        l :& punish \longleftarrow drive, drunk \\
        m :& punish \longleftarrow resist \\
        e :& prison \longleftarrow punish \\
        d :& drive \\
        k :& drunk \\
        r :& resist \\
    \end{aligned}
    \label{eq:example-cg}
    \end{equation}
    \end{minipage}
    \hfill
    \begin{minipage}{0.45\textwidth}
    \centering
    \includegraphics[width=.5\textwidth]{assets/causal_g.pdf}
    \caption{Causal graph of logic program in \cref{eq:example-cg}.}
    \label{fig:causal-g}
    \end{minipage}
\end{figure}

\section{Results}\label{sec:results}

Evaluation was conducted using a publicly available dataset.\footnote{\url{https://www.kaggle.com/datasets/itachi9604/disease-symptom-description-dataset}}
Each record's list of symptoms serves as input.  
The knowledge base—comprising ASP rules automatically derived  
from medical literature—infers potential diagnoses.  
A prediction is considered correct when the disease inferred  
by the ASP solver matches the actual diagnosis listed in the dataset.

\begin{table}[h]
    \centering
    \caption{Accuracy of the framework on selected diseases. 
    The size column indicates the number of logical terms in the disease-specific ASP program.}
    \begin{tabular}{lcc}
    \toprule
    Diseases & Size & Accuracy \\
    \midrule
    Chickenpox     &  66  & 95\% \\
    Pneumonia      &  75  & 100\% \\
    Common Cold    &  44  & 100\% \\
    \bottomrule
    \end{tabular}
\end{table}

High accuracy is achieved when the knowledge base includes a sufficiently 
rich set of logical rules.
% The size of this base, measured as the number of ASP terms associated with a disease, correlates with prediction performance.
In addition to accurate predictions, \sys offers explainability and visualizes the reasoning process behind each diagnosis~(an example is shown in \cref{lst:chickenpox-explanation}).

\begin{lstlisting}[caption={Excerpt from the explanation tree for the chickenpox diagnosis. The structure illustrates the reasoning path based on symptom associations (slightly modified for clarity).},
    label=lst:chickenpox-explanation]{Name}
    *
    |__ diagnosis(chickenpox)
        |__ has(symptom(itching))
        |__ has(symptom(fatigue))
        |__ has(symptom(lethargy))
            |__ has(symptom(fatigue))
            |__ linked_symptom(fatigue, lethargy)
        |__ has(symptom(high_fever))
            |__ has(symptom(mild_fever))
                |__ has(symptom(loss_of_appetite))
                |__ linked_symptom(loss_of_appetite, mild_fever)
            |__ linked_symptom(mild_fever, high_fever)
        |__ has(symptom(loss_of_appetite))
        |__ has(symptom(mild_fever))
            |__ has(symptom(loss_of_appetite))
            |__ linked_symptom(loss_of_appetite, mild_fever)
        |__ has(symptom(swelled_lymph_nodes))
\end{lstlisting}

\Cref{lst:chickenpox-explanation} shows an excerpt from the explanation 
tree generated for a patient diagnosed with \textit{chickenpox}. 
The tree visualizes the reasoning steps performed by the ASP solver to reach this conclusion. 
At the root, \sys infers \texttt{diagnosis(chickenpox)} 
based on the presence of multiple symptoms such as \texttt{itching}, \texttt{fatigue}, \texttt{lethargy}, 
and \texttt{high\_fever}. 
Each symptom is further justified through observed inputs or inferred relationships. 
For example, \texttt{lethargy} is supported by repeated instances of \texttt{fatigue} 
and the known association \texttt{linked\_symptom(fatigue, lethargy)}. 
Similarly, \texttt{high\_fever} is explained via the intermediate symptom \texttt{mild\_fever}, 
which itself is supported by \texttt{loss\_of\_appetite}. 
These explanations demonstrate how \sys combines direct observations 
and transitive symptom relationships encoded in the knowledge base to 
construct a transparent and explainable diagnostic path.

\section{Conclusion}\label{sec:conclusion}

This paper presents \sys, a framework for disease prediction using LLMs and ASP. 
The system uses LLMs to extract knowledge from medical texts and translate it into ASP rules, 
which then drive predictions and provide interpretable explanations. 
The architecture, prompt design, and rule structuring in the knowledge base are described.
Experiments show that \sys can predict diseases accurately 
while keeping the reasoning process transparent.

Future work will focus on scaling this process across more sources, 
formalizing the pipeline to enable auditability and pluggability,
and further experimentation in augmenting each step with additional guardrails, possibly including fine-tuning.
\newpage

\bibliographystyle{unsrt}
\bibliography{bib}

\end{document}